\documentclass[10pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.15em}
\setlist[itemize]{leftmargin=*,itemsep=0.15em,topsep=0.2em,parsep=0em,partopsep=0em}
\newcommand{\sym}[2]{%
  \noindent\begin{minipage}[t]{0.29\textwidth}\raggedright\textbf{\ensuremath{#1}}\end{minipage}%
  \hfill
  \begin{minipage}[t]{0.68\textwidth}\raggedright #2\end{minipage}\par\vspace{0.22em}}
\begin{document}
\small

{\LARGE \textbf{17A Symbol Guide}}\par
{\large Definitions in first-appearance order for Section 17A of the manuscript}\par

\section*{Parameter Manifest}
\begin{itemize}
\item Model family/name: Hubbard-Holstein fixed-manifold McLachlan real-time dynamics, written symbolically for the driven $L=2$ controller as a motivating instance but not restricted to a single calibrated run.
\item Ansatz type(s): an ordered fixed scaffold of Pauli-generator exponential blocks acting on a reference state, with checkpoint-local stay, append, and prune updates of that scaffold.
\item Drive enabled: the guide treats both the static and driven cases symbolically through $H_{\mathrm{stat}}$, $H_{\mathrm{drv}}(t)$, and sampled interval Hamiltonians, with midpoint sampling as the motivating case but left and right sampling also included.
\item Core physical parameters: $t$, $U$, and $dv$ enter implicitly through the coefficients of the static and driven Hamiltonian terms, so the guide names their symbolic containers rather than freezing one numeric physics point.
\item Other run-defining parameters: scaffold length, forecast horizon, calmness thresholds, append and prune thresholds, trust radii, and forecast weights all remain symbolic so the handout stays faithful to the general law written in 17A.
\end{itemize}

\section*{How to read this handout}
Each entry gives the spoken name of the symbol, what kind of mathematical object it is, and the intuition for why it appears, so the list can be used as a compact companion when reading the fully expanded derivations in 17A.

\subsection*{17A.1 Fixed-scaffold projective McLachlan geometry}

\sym{n_k}{The current scaffold length $n_k:=|\mathcal O_k|$ is the number of active ordered generator blocks at checkpoint $k$, so it tells us how many variational coordinates and tangent directions are available before any checkpoint-local structural change is made.}
\sym{\mathcal O_k}{The scaffold $\mathcal O_k$ is the ordered list of currently active generator blocks, so it is the combinatorial object that determines which directions in state space the ansatz is presently allowed to use.}
\sym{d}{The Hilbert-space dimension $d:=\dim\mathcal H$ is the ambient complex vector-space size of the quantum state, so it sets the row dimension of the tangent and drift arrays when they are written componentwise.}
\sym{\mathcal H}{The Hilbert space $\mathcal H$ is the many-body quantum state space on which the ansatz state and Hamiltonian act, so every state vector, tangent vector, and projected drift in this chapter lives inside it.}
\sym{a\text{ in 17A.1}}{The index $a\in\{1,\dots,d\}$ is the Hilbert-space row index used when vectors are written entrywise, so it exists purely to expose matrix and vector formulas in explicit component form.}
\sym{i,j\text{ in 17A.1}}{The indices $i,j\in\{1,\dots,n_k\}$ are scaffold-coordinate indices, so they label tangent columns and coordinate directions in the McLachlan metric, force, and normal equations rather than physical lattice sites.}
\sym{|\psi_k\rangle}{The checkpoint state $|\psi_k\rangle:=|\psi(\theta_k;\mathcal O_k)\rangle$ is the variational many-body state obtained by evaluating the current scaffold at the current coordinate vector, so it is the point on the ansatz manifold whose local geometry is being projected against the sampled Schr"odinger drift.}
\sym{\theta_k}{The coordinate vector $\theta_k$ is the real parameter vector attached to the current scaffold at checkpoint $k$, so it tells us where we are on the fixed-scaffold manifold before the continuous projected motion and before any discrete checkpoint update.}
\sym{U(\theta_k;\mathcal O_k)}{The ordered unitary $U(\theta_k;\mathcal O_k)$ is the scaffold evolution operator built as the product of the active exponential generator blocks, so it converts the static reference state into the current variational state.}
\sym{|\psi_{\mathrm{ref}}\rangle}{The reference state $|\psi_{\mathrm{ref}}\rangle$ is the fixed base vector on which all scaffold unitaries act, so it anchors the entire variational family and gives the manifold a concrete origin in Hilbert space.}
\sym{A_{k,j}}{The generator $A_{k,j}$ is the $j$th active scaffold operator at checkpoint $k$, so it is the primitive Hermitian direction whose exponential block contributes one local variational degree of freedom to the ansatz.}
\sym{U_{k,>j},\,U_{k,<j}}{The partial products $U_{k,>j}$ and $U_{k,<j}$ are the ordered scaffold unitaries to the right and left of the $j$th block, so they isolate the $j$th generator inside the full ordered product and make the tangent derivative formula explicit.}
\sym{\ell}{The index $\ell$ is the running scaffold-block index inside ordered products, so it has no independent physical meaning beyond keeping track of where each exponential block sits relative to the block being differentiated.}
\sym{\widetilde A_{k,j}}{The transported generator $\widetilde A_{k,j}:=U_{k,>j}A_{k,j}U_{k,>j}^{\dagger}$ is the $j$th scaffold generator conjugated by the later blocks in the ordering, so it is the natural operator that acts directly on the current state when the tangent vector is written in local dressed form.}
\sym{{}^{\dagger}}{The dagger operation ${}^{\dagger}$ denotes Hermitian adjoint, so whenever it appears it tells us that an operator or matrix is being turned into its conjugate-transpose counterpart in the standard Hilbert-space inner product.}
\sym{\tau_{k,j}}{The tangent vector $\tau_{k,j}:=\partial_{\theta_{k,j}}|\psi_k\rangle$ is the raw derivative of the variational state with respect to the $j$th scaffold coordinate, so it gives the unprojected local change in the state caused by moving that parameter alone.}
\sym{\bar\tau_{k,j}}{The horizontal tangent $\bar\tau_{k,j}$ is the raw tangent with its component along $|\psi_k\rangle$ removed, so it captures only the physically relevant tangent motion and discards the global-phase direction that would otherwise pollute the geometry.}
\sym{Q_{\psi_k}}{The projector $Q_{\psi_k}:=I-|\psi_k\rangle\langle\psi_k|$ is the orthogonal projector onto the complement of the current state ray, so it strips away the dynamically irrelevant global-phase component from tangent vectors and Hamiltonian drift vectors alike.}
\sym{I\text{ in }Q_{\psi_k}}{The identity operator $I$ in the projector formula is the unit operator on the Hilbert space, so it serves as the ambient neutral map from which the rank-one projector onto the current state is subtracted.}
\sym{\langle\widetilde A_{k,j}\rangle_{\psi_k}}{The expectation value $\langle\widetilde A_{k,j}\rangle_{\psi_k}$ is the current-state mean of the transported generator, so subtracting it centers the generator and converts the tangent into a horizontal covariance-like fluctuation direction.}
\sym{\bar T_k}{The horizontal tangent matrix $\bar T_k$ is the $d\times n_k$ matrix whose $j$th column is $\bar\tau_{k,j}$, so it packages the entire fixed-scaffold tangent plane into one linear map from coordinate velocities to horizontal state-space motion.}
\sym{H_k^{\sharp}}{The sampled interval Hamiltonian $H_k^{\sharp}$ is the piecewise-constant Hamiltonian frozen on checkpoint interval $k$ by the chosen time-sampling rule, so it is the precise drift generator against which the local tangent fit is solved.}
\sym{\bar b_k}{The projected Schr"odinger drift $\bar b_k:=Q_{\psi_k}(-iH_k^{\sharp}|\psi_k\rangle)$ is the horizontal part of the instantaneous Hamiltonian drift, so it is the target vector that the tangent combination tries to match as closely as the current scaffold geometry allows.}
\sym{E_{\psi_k,k}^{\sharp}}{The sampled checkpoint energy $E_{\psi_k,k}^{\sharp}:=\langle\psi_k|H_k^{\sharp}|\psi_k\rangle$ is the expectation value of the frozen interval Hamiltonian in the current state, so it is the scalar that is subtracted off to center the Hamiltonian drift and leave only the fluctuation part.}
\sym{\Re}{The real-part operator $\Re(\cdot)$ extracts the real component of complex overlaps, so it converts complex Hilbert-space inner products into the real geometric quantities that define the McLachlan metric and force on real parameter space.}
\sym{G_k}{The metric matrix $G_k$ is the real covariance matrix of the horizontal tangent columns, so it measures how distinguishable the available tangent directions are and supplies the intrinsic geometry of the current fixed scaffold.}
\sym{f_k}{The force vector $f_k$ is the real overlap of the horizontal tangent matrix with the projected Schr"odinger drift, so it measures how strongly the sampled Hamiltonian wants to move the state in each available variational direction.}
\sym{\Lambda_k}{The regularizer $\Lambda_k$ is a symmetric positive-semidefinite matrix added to the tangent metric, so it stabilizes the local linear solve when the raw geometry is nearly singular or poorly conditioned.}
\sym{K_k}{The regularized normal matrix $K_k:=G_k+\Lambda_k$ is the effective linear system matrix for the checkpoint-local McLachlan solve, so it combines intrinsic tangent geometry with the chosen numerical stabilization.}
\sym{v}{The trial velocity vector $v\in\mathbb R^{n_k}$ is a generic real coordinate-velocity candidate on the current scaffold, so it is the variable over which the local quadratic fit is minimized before the optimal projected velocity is selected.}
\sym{\mathfrak m_k(v)}{The quadratic fit functional $\mathfrak m_k(v)$ is the regularized squared distance between the tangent-generated horizontal motion $\bar T_kv$ and the sampled projected drift $\bar b_k$, so it is the checkpoint-local objective whose minimizer defines the actual McLachlan velocity.}
\sym{\mathcal Q_k(v)}{The projected McLachlan score $\mathcal Q_k(v)$ is the equivalent gain-style quadratic obtained after expanding the fit functional, so it measures how much drift-capture benefit a velocity $v$ achieves before regularization is imposed.}
\sym{\dot\theta_k}{The regularized McLachlan velocity $\dot\theta_k:=K_k^{+}f_k$ is the real coordinate velocity selected on the frozen scaffold at checkpoint $k$, so it is the continuous-time variational motion law used between checkpoint events.}
\sym{{}^{+}}{The superscript ${}^{+}$ denotes Moore-Penrose pseudoinverse, so it allows singular or rank-deficient tangent systems to be solved in the least-squares sense without pretending that every metric block is invertible.}
\sym{\dot\theta_k^{\mathrm{proj}}}{The unregularized projected velocity $\dot\theta_k^{\mathrm{proj}}:=G_k^{+}f_k$ is the pure projective McLachlan solution obtained before adding the regularizer, so it isolates the geometric part of the local fit from the numerical ridge correction.}
\sym{\operatorname{Var}_{\psi_k}(H_k^{\sharp})}{The variance $\operatorname{Var}_{\psi_k}(H_k^{\sharp})=\|\bar b_k\|_2^2$ is the energy-fluctuation scale of the sampled Hamiltonian in the current state, so it provides the intrinsic normalization for the later projection-defect and miss-ratio diagnostics.}

\subsection*{17A.2 Piecewise-constant driven Hamiltonian on each checkpoint interval}

\sym{H_{\mathrm{stat}}}{The static Hamiltonian $H_{\mathrm{stat}}$ is the time-independent Pauli-polynomial part of the physics model, so it carries the baseline hopping, interaction, and other fixed contributions that are present even when the external drive is absent.}
\sym{H_{\mathrm{drv}}(t)}{The driven Hamiltonian $H_{\mathrm{drv}}(t)$ is the explicitly time-dependent Pauli-polynomial contribution, so it is the piece that injects the external control waveform into the real-time dynamics.}
\sym{I_k\text{ in 17A.2}}{The interval $I_k:=[\tau_k,\tau_{k+1}]$ is the $k$th checkpoint time segment, so it is the piece of physical time on which the scaffold is held fixed and one sampled Hamiltonian is used for the projected variational law.}
\sym{\tau_k,\,\tau_{k+1}}{The checkpoint times $\tau_k$ and $\tau_{k+1}$ are consecutive structural update times, so they delimit the open interval on which continuous McLachlan motion occurs before the next discrete decision is made.}
\sym{\Delta\tau_k}{The interval width $\Delta\tau_k:=\tau_{k+1}-\tau_k$ is the physical time duration of checkpoint segment $k$, so it sets the scale for frozen-Hamiltonian projection, exact-reference stepping, and forecast displacement formulas alike.}
\sym{M_k\text{ in 17A.2}}{The integer $M_k\ge 1$ is the exact-reference micro-step count on interval $k$, so it specifies how finely the reference trajectory resolves the time-dependent Hamiltonian inside one coarse checkpoint segment.}
\sym{\sigma_k}{The sampling selector $\sigma_k\in\{0,\tfrac12,1\}$ chooses left, midpoint, or right time-freezing on interval $k$, so it turns the phrase time-sampling rule into a literal scalar that shifts where the Hamiltonian is sampled inside the interval.}
\sym{H(t)}{The full time-dependent Hamiltonian $H(t)=H_{\mathrm{stat}}+H_{\mathrm{drv}}(t)$ is the instantaneous generator of the physical Schr"odinger evolution, so it is the object whose sampled versions drive both the projected variational law and the exact reference path.}
\sym{\mathcal I_{\mathrm{stat}},\,\mathcal I_{\mathrm{drv}}}{The index sets $\mathcal I_{\mathrm{stat}}$ and $\mathcal I_{\mathrm{drv}}$ enumerate the static and driven Pauli terms, so they expose the Hamiltonian as a coefficient map over primitive Pauli strings rather than as an opaque black-box operator.}
\sym{\alpha,\,\beta}{The symbols $\alpha$ and $\beta$ are term labels running over the static and driven Pauli-term index sets, so they keep the fixed and time-dependent Hamiltonian contributions separated in the coefficient decomposition.}
\sym{h_{\alpha}}{The coefficient $h_{\alpha}$ is the static weight of Pauli term $P_{\alpha}$, so it stores the time-independent strength with which that primitive operator enters the baseline Hamiltonian.}
\sym{c_{\beta}(t)}{The coefficient $c_{\beta}(t)$ is the time-dependent weight of driven Pauli term $P_{\beta}$, so it is the explicit scalar control waveform seen by that primitive operator at physical time $t$.}
\sym{P_{\alpha},\,P_{\beta}}{The operators $P_{\alpha}$ and $P_{\beta}$ are primitive Pauli strings in the static and driven decompositions, so they are the operator-level building blocks from which the sampled Hamiltonians are assembled term by term.}
\sym{t_0}{The base time $t_0$ is the global physical start time used to convert checkpoint-local times into absolute sampling times, so it anchors every sampled Hamiltonian to a single laboratory time origin.}
\sym{t_k^{\sharp}}{The sampled checkpoint time $t_k^{\sharp}:=t_0+\tau_k+\sigma_k\Delta\tau_k$ is the absolute time at which the interval Hamiltonian is frozen for the projected variational solve, so it turns the abstract sampling rule into a concrete time argument of the driven coefficients.}
\sym{\delta\tau_k}{The micro-step size $\delta\tau_k:=\Delta\tau_k/M_k$ is the refined exact-reference time increment on interval $k$, so it controls how the exact driven propagator resolves the coarse checkpoint interval.}
\sym{\mu}{The index $\mu$ is the micro-step counter inside interval $k$, so it labels the refined subintervals on which the exact reference path samples the driven Hamiltonian.}
\sym{t_{k,\mu}^{\sharp}}{The micro-sampled time $t_{k,\mu}^{\sharp}:=t_0+\tau_k+(\mu+\sigma_k)\delta\tau_k$ is the absolute time at which the exact reference propagator freezes the Hamiltonian on micro-step $\mu$, so it mirrors the same left-midpoint-right doctrine used by the coarse projected law.}
\sym{H_{k,\mu}^{\sharp}}{The micro-sampled Hamiltonian $H_{k,\mu}^{\sharp}:=H(t_{k,\mu}^{\sharp})$ is the exact-reference Hamiltonian used on refined substep $\mu$, so it is the fine-grained counterpart of the coarse sampled Hamiltonian $H_k^{\sharp}$.}
\sym{|\psi_{k,\mu}^{\mathrm{ex}}\rangle}{The exact-reference micro-step state $|\psi_{k,\mu}^{\mathrm{ex}}\rangle$ is the state reached after $\mu$ refined propagations on interval $k$, so it gives the benchmark path against which later forecast errors and checkpoint decisions are judged.}

\subsection*{17A.3 Residual geometry and local inadequacy of the scaffold}

\sym{\bar r_k}{The residual $\bar r_k:=\bar T_k\dot\theta_k-\bar b_k$ is the horizontal drift mismatch left over after solving the regularized McLachlan problem on the current scaffold, so it is the immediate geometric witness that the available tangent plane may still be too small.}
\sym{\epsilon_{\mathrm{step},k}}{The step defect $\epsilon_{\mathrm{step},k}$ is the norm of the realized residual after the regularized solve, so it measures the actual mismatch incurred by the chosen numerical velocity rather than the ideal mismatch of the unregularized projective optimum.}
\sym{\epsilon_{\mathrm{proj},k}}{The projective defect $\epsilon_{\mathrm{proj},k}$ is the minimum possible horizontal mismatch on the current scaffold before regularization is considered, so it isolates the structural inadequacy of the tangent space itself from the additional distortion introduced by numerical stabilization.}
\sym{\rho_{\mathrm{miss},k}}{The miss ratio $\rho_{\mathrm{miss},k}$ is the projective defect normalized by the current Hamiltonian variance, so it is the scale-free checkpoint statistic that says how much of the physically relevant drift the current scaffold is failing to capture.}
\sym{\varepsilon}{The small positive scalar $\varepsilon$ is a denominator regularization used in normalized ratios, so it prevents division-by-zero pathologies when the instantaneous Hamiltonian variance becomes extremely small.}

\subsection*{17A.4 Checkpoint-local decision surfaces and calmness gate}

\sym{x,\,y\text{ in 17A.4}}{The vectors $x\in\mathbb R^m$ and $y\in\mathbb R^n$ are two coordinate vectors of possibly different lengths that are being compared across scaffold changes, so they provide the raw material for a distance notion that respects zero-padding and overlap when the parameter dimension changes.}
\sym{m,\,n\text{ in 17A.4}}{The integers $m$ and $n$ are the coordinate lengths of the two vectors being compared in the overlap norm, so they exist only to say how much of each vector must be matched directly and how much must be treated as unmatched tail coordinates.}
\sym{r\text{ in 17A.4}}{The symbol $r:=\min(m,n)$ is the overlap length of two differently sized coordinate vectors in the checkpoint comparison norm, so here it means shared coordinate extent rather than an append record.}
\sym{\|x-y\|_{\mathrm{ov}}}{The overlap norm $\|x-y\|_{\mathrm{ov}}$ is the zero-padded Euclidean distance between coordinate vectors of unequal length, so it lets the controller compare successive velocities fairly even when a checkpoint action changes the scaffold dimension.}
\sym{\dot\theta_{k-1}^{\mathrm{acc}}}{The previously accepted velocity $\dot\theta_{k-1}^{\mathrm{acc}}$ is the last checkpoint velocity after alignment to the current coordinate dimension, so it is the natural reference direction against which calmness and directional continuity are measured.}
\sym{c_k^{\mathrm{dir}}}{The directional calmness score $c_k^{\mathrm{dir}}$ is the cosine similarity between the current and previously accepted velocities, so it detects whether local motion is continuing in a stable direction or attempting to turn sharply.}
\sym{\eta_k^{\mathrm{rate}}}{The rate-change score $\eta_k^{\mathrm{rate}}$ is the normalized overlap distance between consecutive accepted velocities, so it measures how abruptly the control law itself is changing from one checkpoint to the next.}
\sym{\mathfrak C_k}{The calmness gate $\mathfrak C_k$ is the binary indicator that both directional similarity and normalized rate change pass their thresholds, so it marks the regime in which deletion is allowed because the scaffold is moving quietly enough to be pruned safely.}
\sym{\tau_{\mathrm{calm}}^{\cos},\,\tau_{\mathrm{calm}}^{\mathrm{rate}},\,\tau_{\mathrm{miss}},\,\tau_{\mathrm{pr}}}{The thresholds $\tau_{\mathrm{calm}}^{\cos}$, $\tau_{\mathrm{calm}}^{\mathrm{rate}}$, $\tau_{\mathrm{miss}}$, and $\tau_{\mathrm{pr}}$ are the checkpoint calmness, miss-dominance, and prune-entry gates, so together they partition the controller into the large-miss forecast regime, the calm low-miss prune regime, and the fallback stay regime.}
\sym{\mathcal P_k^{\mathrm{stay}}}{The set $\mathcal P_k^{\mathrm{stay}}$ is the finite family of normalized same-scaffold proposal directions considered by stay actions, so it turns the stay choice into a forecast comparison over a discrete menu of geometric motions rather than one hard-coded velocity.}
\sym{\mathcal A_k}{The set $\mathcal A_k$ is the finite discrete step-scale set used by stay and append actions, so it controls how aggressively a proposal direction is extrapolated into the short forecast horizon.}
\sym{\mathcal R_k^{\mathrm{conf}}}{The set $\mathcal R_k^{\mathrm{conf}}$ is the shortlist of append records that survive exact or compressed confirm scoring, so it is the admissible append menu presented to the miss-dominant forecast law.}
\sym{\mathcal U_k^{\mathrm{stay}},\,\mathcal U_k^{\mathrm{app}}}{The sets $\mathcal U_k^{\mathrm{stay}}$ and $\mathcal U_k^{\mathrm{app}}$ are the admissible stay and append action classes built from proposal directions or confirmed append records together with discrete scales, so they are the actual action spaces over which the forecast score is minimized.}
\sym{\nu_k^{\star}\text{ in 17A.4}}{The checkpoint choice $\nu_k^{\star}$ is the action selected at checkpoint $k$ after applying the miss and calmness gates, so it is the discrete bridge between continuous projected motion and structural scaffold change.}

\subsection*{17A.5 Position-aware append geometry and confirm-grade gain}

\sym{r=(m,p)\text{ in 17A.5}}{The append record $r=(m,p)$ means insert candidate block $m$ at ordered position $p$ inside the current scaffold, so here the letter $r$ no longer means overlap length and instead names one concrete scaffold-growth move.}
\sym{m_r}{The integer $m_r$ is the number of new coordinates carried by append record $r$, so it determines the width of the new tangent block and the size of the new-coordinate Schur complement.}
\sym{\eta\text{ in 17A.5}}{The vector $\eta=(\eta_1,\dots,\eta_{m_r})$ collects the new coordinates created by an append move, so it parameterizes motion in the newly added tangent directions while the old coordinates remain fixed at the checkpoint instant.}
\sym{\vartheta}{The augmented coordinate vector $\vartheta=(\theta,\eta)$ concatenates the old scaffold coordinates and the new append coordinates, so it is the natural parameter vector on the enlarged manifold after a candidate block has been inserted.}
\sym{\bar U_{r,k}}{The appended tangent matrix $\bar U_{r,k}$ is the horizontal tangent block generated by differentiating the augmented state with respect to the new coordinates at zero initialization, so it captures exactly which new Hilbert-space directions would become available if record $r$ were appended.}
\sym{s\text{ in 17A.5}}{The index $s\in\{1,\dots,d\}$ is the Hilbert-space row index used when writing the appended tangent block entrywise, so it plays the same component-exposure role here that $a$ played for the original tangent matrix in 17A.1.}
\sym{a\text{ in 17A.5}}{The index $a\in\{1,\dots,m_r\}$ now labels one of the new append coordinates rather than a Hilbert-space row, so it is an intentional overload that tracks the columns of the appended tangent block and the entries of the new-coordinate force vector.}
\sym{B_{r,k}}{The coupling block $B_{r,k}$ is the real overlap matrix between old tangent columns and newly appended tangent columns, so it measures how strongly the candidate append directions interact with the tangent geometry that already exists on the scaffold.}
\sym{C_{r,k}}{The self block $C_{r,k}$ is the real overlap matrix of the appended tangent block with itself, so it is the intrinsic metric of the new tangent coordinates before any elimination against the old coordinates is performed.}
\sym{q_{r,k}}{The appended force vector $q_{r,k}$ is the real overlap of the appended tangent block with the projected drift, so it measures how strongly the sampled Hamiltonian wants to move the state directly into the candidate new directions.}
\sym{\xi}{The vector $\xi\in\mathbb R^{m_r}$ is a generic new-coordinate velocity for the appended block, so it is the Schur-complement variable that remains after the old coordinates are eliminated exactly.}
\sym{\Lambda_r}{The regularizer $\Lambda_r$ is the positive-semidefinite regularization applied to the new append coordinates, so it stabilizes the appended Schur system in the same way that $\Lambda_k$ stabilizes the old tangent solve.}
\sym{\mathfrak m_{r,k}(v,\xi)}{The augmented fit functional $\mathfrak m_{r,k}(v,\xi)$ is the regularized squared mismatch between the projected drift and a tangent combination using both old and candidate new directions, so it is the local append trial objective whose elimination yields the exact append gain.}
\sym{\dot\theta_{r,k}^{\mathrm{old}},\,\dot\eta_{r,k}}{The vectors $\dot\theta_{r,k}^{\mathrm{old}}$ and $\dot\eta_{r,k}$ are the old-coordinate and new-coordinate components of the augmented McLachlan solve for append record $r$, so together they describe the best local velocity on the enlarged scaffold at the checkpoint instant.}
\sym{v^{\star}(\xi)}{The reduced old-coordinate optimizer $v^{\star}(\xi)$ is the exact old-block minimizer obtained after fixing the new-coordinate trial vector $\xi$, so it is the algebraic step that turns the augmented least-squares problem into a pure Schur-complement problem in the append coordinates alone.}
\sym{w_{r,k}}{The residualized append force $w_{r,k}$ is the new-coordinate force left over after projecting out what the old scaffold can already explain, so it is the genuine incremental drift demand that only the candidate append block can satisfy.}
\sym{S_{r,k}}{The Schur complement $S_{r,k}$ is the effective new-coordinate metric after exact elimination of the old coordinates, so it packages the true incremental geometry of the append move into the matrix whose pseudoinverse controls the exact gain.}
\sym{\Delta_{r,k}^{\mathrm{exact}}}{The exact append gain $\Delta_{r,k}^{\mathrm{exact}}$ is the Schur-complement quadratic improvement achieved by adding record $r$, so it is the local geometric quantity that tells us how much the current drift fit would improve if the scaffold were enlarged in precisely that way.}
\sym{\rho_{r,k}^{\mathrm{gain}}}{The normalized append gain $\rho_{r,k}^{\mathrm{gain}}$ is the exact append gain divided by the instantaneous Hamiltonian variance scale, so it converts raw Schur improvement into a size-free checkpoint score that can be compared across states and times.}
\sym{V_k,\,\Sigma_k}{The matrices $V_k$ and $\Sigma_k$ are the supported spectral factors in $K_k=V_k\Sigma_k^2V_k^{\top}$, so they expose the old-coordinate geometry in whitened modal coordinates where compressed confirm scoring becomes algebraically transparent.}
\sym{z_k}{The whitened force $z_k:=\Sigma_k^{-1}V_k^{\top}f_k$ is the old-coordinate force vector expressed in supported metric-normal coordinates, so it is the baseline drift demand seen from the geometry in which confirm compression is easiest to interpret.}
\sym{\Gamma_{r,k}}{The whitened coupling matrix $\Gamma_{r,k}:=\Sigma_k^{-1}V_k^{\top}B_{r,k}$ is the old-new interaction block expressed in supported metric-normal coordinates, so it tells confirm scoring which old modes the append block actually couples to once geometry scaling has been stripped away.}
\sym{J_r}{The retained mode set $J_r$ is the subset of whitened old modes kept during compressed confirm scoring, so it turns confirm into a deliberately reduced measurement problem that can still lower-bound the exact append gain.}
\sym{\widetilde S_{r,k}(J_r)}{The compressed Schur block $\widetilde S_{r,k}(J_r)$ is the Schur-complement approximation obtained by retaining only the old modes in $J_r$, so it is the confirm-stage surrogate metric used when one wants a cheaper but still structured gain estimate.}
\sym{\widetilde\Delta_{r,k}(J_r)}{The compressed append gain $\widetilde\Delta_{r,k}(J_r)$ is the Schur improvement computed from the compressed Schur block, so it is the confirm-stage lower-cost proxy for the exact append gain.}
\sym{\dot\vartheta_{r,k}^{\mathrm{aug}}}{The augmented append velocity $\dot\vartheta_{r,k}^{\mathrm{aug}}$ is the combined old-and-new coordinate velocity proposed on the enlarged scaffold for record $r$, so it is the object whose direction is compared against the previously accepted checkpoint motion in the confirm penalty.}
\sym{\delta_{r,k}^{\mathrm{dir}}}{The append direction penalty $\delta_{r,k}^{\mathrm{dir}}$ is the overlap-norm distance between the augmented append velocity and the previously accepted checkpoint velocity, so it discourages candidate append moves that improve the drift fit only by violently reorienting the motion law.}
\sym{g_{r,k}^{\mathrm{new}}}{The quantity $g_{r,k}^{\mathrm{new}}$ is the measurement or acquisition burden assigned to append record $r$, so it lets confirm scoring trade geometric improvement against the additional observational cost of activating that candidate.}
\sym{\lambda_{\mathrm{dir}},\,\lambda_{\mathrm{meas}}}{The weights $\lambda_{\mathrm{dir}}$ and $\lambda_{\mathrm{meas}}$ penalize directional disruption and new-measurement burden inside the confirm score, so they tune how aggressively the controller prefers stable and cheap append moves over merely high-gain ones.}
\sym{S_{r,k}^{\mathrm{conf}}}{The confirm score $S_{r,k}^{\mathrm{conf}}$ is the penalized exact-or-compressed append merit assigned to record $r$, so it is the final shortlist score that decides which append records are even allowed into the forecast comparison.}
\sym{\tau_{\mathrm{gain}},\,\tau_{\mathrm{conf}}}{The thresholds $\tau_{\mathrm{gain}}$ and $\tau_{\mathrm{conf}}$ are the gain and confirm gates for append records, so they prevent weak or unstable scaffold-growth moves from entering the action set even before the forecast law compares them against stay.}

\subsection*{17A.6 Proposal family for within-scaffold motion}

\sym{M_k\text{ in 17A.6}}{The symbol $M_k$ now denotes the proposal metric used to normalize and compare same-scaffold forecast directions, so this is a deliberate overload of the earlier micro-step count notation and here it means a geometry for proposal magnitudes rather than a time-refinement integer.}
\sym{\|v\|_{M_k},\,\langle x,y\rangle_{M_k}}{The norm $\|v\|_{M_k}$ and inner product $\langle x,y\rangle_{M_k}$ are the metric-induced magnitude and angle notions associated with the proposal metric, so they let stay proposals be compared and blended in a way that respects the local geometry rather than raw Euclidean coordinates.}
\sym{I\text{ in }M_k=I}{The identity matrix $I$ appearing in the fallback definition of the proposal metric is the coordinate-space identity, so it provides a neutral Euclidean metric when one does not want to normalize proposals by the tangent geometry.}
\sym{I_k^{\mathrm{drv}}}{The index block $I_k^{\mathrm{drv}}$ is the subset of scaffold coordinates regarded as drive-aligned at checkpoint $k$, so it marks the coordinate directions on which drive-only proposal components are allowed to live.}
\sym{\mathcal B_k}{The finite set $\mathcal B_k$ is the blend-weight menu for same-scaffold proposal mixing, so it parameterizes how much orthogonalized drive-aware content is injected into the baseline McLachlan direction.}
\sym{v_k^{\mathrm{McL}}}{The baseline proposal $v_k^{\mathrm{McL}}$ is simply the checkpoint McLachlan velocity itself, so it represents the default same-scaffold direction before any drive-aware or secant-aware proposal engineering is added.}
\sym{v_{k,\mathrm{drv}}^{\mathrm{cur}}}{The current drive-restricted proposal $v_{k,\mathrm{drv}}^{\mathrm{cur}}$ is the velocity obtained by solving the regularized system only on the drive-aligned coordinate block using current-time data, so it isolates the tangent motion that the present drive signal alone is asking for.}
\sym{v_{k,\mathrm{drv}}^{\mathrm{next}}}{The lookahead drive-restricted proposal $v_{k,\mathrm{drv}}^{\mathrm{next}}$ is the corresponding block-restricted solve using next-checkpoint sampled geometry, so it injects a one-step anticipatory response to the near-future drive profile into the stay proposal family.}
\sym{K_{k+1}^{\sharp},\,f_{k+1}^{\sharp}}{The objects $K_{k+1}^{\sharp}$ and $f_{k+1}^{\sharp}$ are the next-sampled regularized geometry and force restricted to the drive-aligned block, so they give the lookahead drive proposal a concrete future tangent system to solve against.}
\sym{d_k}{The direction $d_k$ is the generic auxiliary drive-aware direction that is orthogonalized against the baseline McLachlan motion, so it is the raw ingredient from which norm-locked blended proposals are built.}
\sym{d_k^{\perp}}{The orthogonalized direction $d_k^{\perp}$ is the component of the auxiliary direction that is $M_k$-orthogonal to the baseline McLachlan proposal, so it isolates genuinely new proposal content rather than merely rescaling what the baseline already does.}
\sym{\widetilde d_k^{\perp}}{The norm-locked orthogonalized direction $\widetilde d_k^{\perp}$ is the orthogonal auxiliary direction rescaled to have the same $M_k$-norm as the baseline McLachlan proposal, so it makes blend weights comparable without changing the overall proposal size scale.}
\sym{\beta}{The scalar $\beta\in\mathcal B_k$ is the blend weight that determines how much orthogonalized auxiliary content is mixed into the baseline McLachlan proposal, so it parametrizes a small finite family of same-scaffold forecast directions.}
\sym{v_{k,\beta}^{\mathrm{blend}}}{The norm-locked blend $v_{k,\beta}^{\mathrm{blend}}$ is the rescaled mixture of the baseline McLachlan proposal and the orthogonalized auxiliary direction, so it explores drive-aware deviations from the local McLachlan step without allowing uncontrolled changes in overall proposal magnitude.}
\sym{\widehat v}{The normalized proposal $\widehat v:=v/\|v\|_{M_k}$ is the metric-unit version of any nonzero proposal direction, so it separates direction choice from scale choice and makes the forecast law compare geometry and step size as two distinct ingredients.}
\sym{\widehat v_k^{\mathrm{sec}}}{The symbol $\widehat v_k^{\mathrm{sec}}$ denotes the normalized secant-based proposal imported from 17A.7, so it is the same-scaffold direction obtained by projecting a short exact secant displacement back into the current tangent coordinates.}

\subsection*{17A.7 Tangent-secant forecast proposal and signed-energy-lead taper}

\sym{R_k}{The trust radius $R_k$ is the maximum allowed $M_k$-metric size of the secant-derived proposal before tapering, so it prevents an exact-reference secant from producing an implausibly large tangent step on one checkpoint interval.}
\sym{\eta_E}{The scalar $\eta_E$ is the relative energy-lead cap used in the secant taper, so it limits how much the secant proposal is allowed to run ahead of the exact energy trend before being suppressed.}
\sym{|\widetilde\psi_{k+1}^{\mathrm{ex}}\rangle}{The phase-aligned exact state $|\widetilde\psi_{k+1}^{\mathrm{ex}}\rangle$ is the next exact reference state after removing the arbitrary global phase relative to the current state, so it turns the exact one-step difference into a geometrically meaningful secant displacement.}
\sym{\delta_k^{\mathrm{sec}}}{The projected secant displacement $\delta_k^{\mathrm{sec}}$ is the horizontal component of the phase-aligned exact one-step state difference, so it captures what the exact trajectory is trying to do over one interval after both global phase and pure ray-parallel motion have been removed.}
\sym{\xi_k^{\mathrm{sec}}}{The tangent secant coordinates $\xi_k^{\mathrm{sec}}$ are the real overlaps of the projected secant displacement with the current tangent matrix, so they express the exact one-step reference motion in local scaffold coordinates before time rescaling.}
\sym{v_k^{\mathrm{sec}}}{The secant proposal velocity $v_k^{\mathrm{sec}}$ is the tangent-coordinate inverse image of the projected exact secant divided by the interval length and then trust clipped, so it is the stay proposal that most directly mimics short-horizon exact motion while still living on the current scaffold.}
\sym{b_k^{E}}{The energy bias $b_k^{E}:=E_k^{\mathrm{ctrl}}-E_k^{\mathrm{ex}}$ is the signed controller-versus-exact energy difference at checkpoint $k$, so it measures whether the present variational trajectory is already leading or lagging the exact energy profile.}
\sym{E_k^{\mathrm{ctrl}},\,E_k^{\mathrm{ex}}}{The quantities $E_k^{\mathrm{ctrl}}$ and $E_k^{\mathrm{ex}}$ are the controller-side and exact sampled energies at checkpoint $k$, so they give the secant taper and later forecast diagnostics a common energy reference at the current checkpoint.}
\sym{\Delta E_k^{\mathrm{ex}}}{The exact energy increment $\Delta E_k^{\mathrm{ex}}:=E_{k+1}^{\mathrm{ex}}-E_k^{\mathrm{ex}}$ is the one-interval energy change along the exact reference path, so it tells the taper which energy direction the exact dynamics is presently moving in.}
\sym{\ell_k^{E}}{The signed-energy-lead scalar $\ell_k^{E}:=[\operatorname{sgn}(\Delta E_k^{\mathrm{ex}})b_k^{E}]_{+}$ is the nonnegative amount by which the controller is already ahead of the exact energy trend, so it measures when the secant proposal should be damped rather than trusted.}
\sym{\chi_k^{E}}{The taper factor $\chi_k^{E}$ is the multiplicative suppression applied to the secant proposal when the controller energy is leading the exact energy by too much, so it protects the secant direction from becoming an overaggressive extrapolation.}

\subsection*{17A.8 Exact-forecast horizon score and stay-append decision law}

\sym{h}{The horizon index $h\in\{1,\dots,H_k\}$ labels one future checkpoint inside the forecast window, so it tells the score which predicted future state or observable sample is being compared against the exact reference at that stage.}
\sym{H_k\text{ in 17A.8}}{The horizon length $H_k$ is the number of future checkpoints included in the forecast score at checkpoint $k$, so it controls how farsighted the stay-versus-append comparison is allowed to be.}
\sym{\omega_h}{The positive weight $\omega_h$ is the importance assigned to horizon step $h$, so it lets the forecast law emphasize near-term or farther-term tracking errors according to the chosen checkpoint-control philosophy.}
\sym{W_k}{The normalizing weight sum $W_k:=\sum_{h=1}^{H_k}\omega_h$ is the total horizon mass used to average forecast errors, so it converts the weighted forecast score into a normalized quantity rather than an unscaled sum.}
\sym{u}{The symbol $u$ denotes a generic checkpoint action in the forecast comparison, so it stands for either a stay proposal plus scale or an append record plus scale while the score is being minimized.}
\sym{p\text{ in 17A.8}}{The symbol $p$ inside a stay action is the normalized same-scaffold proposal direction chosen from $\mathcal P_k^{\mathrm{stay}}$, so here it means a forecast direction rather than an insertion position.}
\sym{a\text{ in 17A.8}}{The symbol $a$ inside an action is the discrete forecast scale chosen from $\mathcal A_k$, so here it means how far a stay or append proposal is extrapolated over each checkpoint interval rather than any matrix row or append-coordinate index.}
\sym{p_{\mathrm{ins}}}{The insertion position $p_{\mathrm{ins}}$ is the ordered slot at which an append record inserts its new block into the scaffold, so it distinguishes append actions that use the same candidate block but place it at different points in the ordered ansatz.}
\sym{\lambda_F,\,\lambda_s,\,\lambda_D,\,\lambda_n,\,\lambda_E,\,\lambda_{\mathrm{slope}},\,\lambda_{\mathrm{curv}},\,\lambda_{\mathrm{under}},\,\lambda_{\mathrm{over}}}{These nonnegative weights are the forecast-score penalties attached to fidelity, staggered density, doublon, site occupation, total energy, slope mismatch, curvature mismatch, under-response, and over-response, so together they encode which kinds of tracking errors the checkpoint controller is willing to pay for and which ones it tries hardest to avoid.}
\sym{\theta_{k+h\mid k}^{(\mathrm{stay},p,a)}}{The predicted stay coordinates $\theta_{k+h\mid k}^{(\mathrm{stay},p,a)}$ are the checkpoint-$k$ stay forecast coordinates at horizon step $h$, so they are the extrapolated same-scaffold parameters used to generate predicted future states under a chosen stay action.}
\sym{|\psi_{k+h}^{(\mathrm{stay},p,a)}\rangle}{The predicted stay state $|\psi_{k+h}^{(\mathrm{stay},p,a)}\rangle$ is the future variational state obtained by evaluating the current scaffold at the stay-forecast coordinates, so it is the state whose observables are compared against the exact reference when scoring a stay action.}
\sym{\vartheta_k^{(r)}}{The zero-lifted augmented coordinates $\vartheta_k^{(r)}:=(\theta_k,0)$ are the enlarged-scaffold coordinates at the append instant before the new block has moved, so they preserve the old state exactly while exposing the new tangent directions of append record $r$.}
\sym{\vartheta_{k+h\mid k}^{(\mathrm{append},r,a)}}{The predicted append coordinates $\vartheta_{k+h\mid k}^{(\mathrm{append},r,a)}$ are the augmented-scaffold coordinates reached by extrapolating the append velocity over $h$ future intervals, so they generate the forecast states used to score an append action against exact dynamics.}
\sym{|\psi_{k+h}^{(\mathrm{append},r,a)}\rangle}{The predicted append state $|\psi_{k+h}^{(\mathrm{append},r,a)}\rangle$ is the future variational state on the enlarged scaffold produced by record $r$, so it is the append counterpart of the predicted stay state in the forecast comparison.}
\sym{|\psi_{k+h}^{\mathrm{ex}}\rangle}{The exact reference state $|\psi_{k+h}^{\mathrm{ex}}\rangle$ is the exact driven state at future checkpoint step $h$, so it is the benchmark against which every forecast observable and overlap error is measured.}
\sym{|\phi\rangle,\,\phi_x,\,|x\rangle}{The generic state $|\phi\rangle=\sum_x\phi_x|x\rangle$ is a basis expansion used to define forecast observables explicitly, where $|x\rangle$ is a computational many-body basis state and $\phi_x$ is its complex amplitude, so these symbols turn occupation and doublon observables into transparent weighted sums over basis configurations.}
\sym{j\text{ in 17A.8}}{The index $j\in\{1,\dots,L\}$ is now a physical lattice-site index rather than a scaffold-coordinate index, so from this point onward it labels where particles or doublons live in real space when observables are evaluated.}
\sym{\sigma\text{ in 17A.8}}{The spin label $\sigma\in\{\uparrow,\downarrow\}$ distinguishes the up-spin and down-spin occupancy channels at a given lattice site, so it is needed to define both site densities and the on-site doublon observable.}
\sym{b_{j,\sigma}(x)}{The bit $b_{j,\sigma}(x)\in\{0,1\}$ is the occupation value of basis configuration $x$ at site $j$ and spin $\sigma$, so it turns occupation observables into explicit basis-level weighted sums over probabilities $|\phi_x|^2$.}
\sym{\hat n_j}{The site-density operator $\hat n_j:=\hat n_{j,\uparrow}+\hat n_{j,\downarrow}$ is the total fermion number on lattice site $j$, so it is the local observable from which both raw site occupations and the staggered density pattern are built.}
\sym{n_{j,\sigma}[\phi],\,n_j[\phi]}{The quantities $n_{j,\sigma}[\phi]$ and $n_j[\phi]$ are the spin-resolved and total site occupations of state $|\phi\rangle$, so they are the forecast observables that tell the controller how well a candidate action reproduces local density physics on each lattice site.}
\sym{\hat D,\,D[\phi]}{The operator $\hat D:=\sum_{j=1}^{L}\hat n_{j,\uparrow}\hat n_{j,\downarrow}$ and its expectation $D[\phi]$ are the total doublon count and doublon expectation, so they measure how much on-site double occupancy a forecast trajectory creates or suppresses relative to the exact reference.}
\sym{\hat S,\,s[\phi]}{The operator $\hat S:=\frac{1}{L}\sum_{j=1}^{L}(-1)^{j-1}\hat n_j$ and its expectation $s[\phi]$ are the staggered density observable and its value, so they compress the alternating spatial density pattern into one scalar that is especially useful for driven and dimerized responses.}
\sym{E_{k+h}^{\mathrm{ctrl}}(u),\,E_{k+h}^{\mathrm{ex}},\,H_{k+h}^{\sharp}}{The quantities $E_{k+h}^{\mathrm{ctrl}}(u)$ and $E_{k+h}^{\mathrm{ex}}$ are the controller-side and exact energies evaluated with the sampled future Hamiltonian $H_{k+h}^{\sharp}$, so they provide the forecast score with a common energy language for comparing predicted variational motion to the exact reference path.}
\sym{\varepsilon_{k+h}^{F}(u),\,\varepsilon_{k+h}^{s}(u),\,\varepsilon_{k+h}^{D}(u),\,\varepsilon_{k+h}^{n}(u),\,\varepsilon_{k+h}^{E}(u)}{These five discrepancy terms are the horizon-$h$ fidelity, staggered-density, doublon, site-occupation, and energy errors of action $u$, so they are the first-order forecast observables whose weighted average makes up the basic tracking part of the checkpoint decision score.}
\sym{\varpi_h^{(1)},\,\varpi_h^{(2)}}{The averaged weights $\varpi_h^{(1)}$ and $\varpi_h^{(2)}$ are the adjacent-step and three-point weight averages used in slope and curvature penalties, so they let discrete energy-derivative mismatches be weighted compatibly with the main horizon score.}
\sym{\varrho_E}{The relative excursion band $\varrho_E\in[0,1)$ sets the tolerated fractional deviation between forecast and exact signed energy excursions, so it defines what counts as under-response or over-response in the energy-shape penalties.}
\sym{\Delta E_{k+h}^{\mathrm{ctrl}}(u),\,\Delta E_{k+h}^{\mathrm{ex}}}{These forward differences are the one-step controller-side and exact energy increments along the forecast horizon, so they measure whether a candidate action gets the local trend of the energy curve correct rather than merely matching isolated energy values.}
\sym{\Delta^2E_{k+h}^{\mathrm{ctrl}}(u),\,\Delta^2E_{k+h}^{\mathrm{ex}}}{These second forward differences are the controller-side and exact energy curvatures along the forecast horizon, so they measure whether a candidate action reproduces the bending or shape of the exact energy trajectory over several future checkpoints.}
\sym{\Xi_k^{\mathrm{shape,slope}}(u),\,\Xi_k^{\mathrm{shape,curv}}(u)}{The shape penalties $\Xi_k^{\mathrm{shape,slope}}(u)$ and $\Xi_k^{\mathrm{shape,curv}}(u)$ are weighted averages of slope and curvature mismatches, so they punish forecast actions that get the local trend or local bending of the energy path wrong even when pointwise errors look acceptable.}
\sym{x_{k+h}^{\mathrm{ctrl}}(u),\,x_{k+h}^{\mathrm{ex}}}{The signed excursion variables $x_{k+h}^{\mathrm{ctrl}}(u)$ and $x_{k+h}^{\mathrm{ex}}$ are the forecast and exact energy displacements from their current checkpoint baselines, so they let the score compare the size and direction of predicted energy excursions relative to the current state rather than in absolute-energy coordinates alone.}
\sym{\Pi_{k+h}^{\mathrm{ctrl}}(u)}{The signed forecast excursion $\Pi_{k+h}^{\mathrm{ctrl}}(u)$ is the controller energy displacement projected onto the sign of the exact excursion, so it tells the score whether the candidate action is responding in the correct energetic direction and by how much.}
\sym{L_{k+h}^{\mathrm{ex}},\,U_{k+h}^{\mathrm{ex}}}{The bounds $L_{k+h}^{\mathrm{ex}}$ and $U_{k+h}^{\mathrm{ex}}$ are the lower and upper acceptable exact-excursion bands derived from $\varrho_E$, so they determine when a forecast under-responds or over-responds relative to the exact energy motion.}
\sym{\Xi_k^{\mathrm{exc,under}}(u),\,\Xi_k^{\mathrm{exc,over}}(u)}{The excursion penalties $\Xi_k^{\mathrm{exc,under}}(u)$ and $\Xi_k^{\mathrm{exc,over}}(u)$ measure how far the signed forecast excursion falls below or rises above the acceptable exact band, so they explicitly reward energetic responsiveness that is neither too weak nor too aggressive.}
\sym{\mathcal J_k^{\mathrm{forecast}}(u)}{The forecast score $\mathcal J_k^{\mathrm{forecast}}(u)$ is the full weighted horizon functional combining overlap, observable, energy, slope, curvature, and excursion penalties for action $u$, so it is the definitive scalar objective minimized when the miss ratio is large enough to open stay-versus-append competition.}
\sym{\mathcal J_k^{\mathrm{trk}}(u)}{The shorthand $\mathcal J_k^{\mathrm{trk}}(u)$ denotes the pure first-order tracking part of the forecast score before shape and excursion corrections are added, so it separates raw observable fidelity from the more refined energetic-response penalties.}
\sym{\nu_{k,\mathrm{stay}}^{\star},\,\nu_{k,\mathrm{app}}^{\star}}{The actions $\nu_{k,\mathrm{stay}}^{\star}$ and $\nu_{k,\mathrm{app}}^{\star}$ are the best stay and best append candidates under the full horizon score, so they are the two finalists that the checkpoint law compares before deciding whether the scaffold actually grows.}
\sym{\Delta F_k}{The scalar $\Delta F_k$ is the one-step fidelity advantage of the best append candidate over the best stay candidate, so it explicitly records whether growing the scaffold helps or hurts immediate exact-state tracking.}
\sym{\Delta\varepsilon_k^{E}}{The scalar $\Delta\varepsilon_k^{E}$ is the one-step energy-error difference between the best append and best stay candidates, so it lets the checkpoint law veto append moves that improve fidelity only by making the energetic forecast materially worse.}
\sym{\tau_F,\,\tau_E}{The thresholds $\tau_F$ and $\tau_E$ are the fidelity-advantage and energy-veto gates in the final stay-versus-append choice, so they encode how much immediate tracking benefit an append action must show before it is allowed to beat the best stay action.}

\subsection*{17A.9 Exact local prune law in the calm regime}

\sym{j\text{ in 17A.9}}{The symbol $j$ now denotes a logical scaffold block index in the prune law rather than a lattice site or scaffold-coordinate column, so it identifies which existing scaffold block is being considered for deletion in the calm regime.}
\sym{I_j,\,I_j^{c}}{The set $I_j$ is the runtime coordinate block occupied by scaffold block $j$ and $I_j^{c}$ is its complement, so these sets identify exactly which coordinates are removed and which are retained when the prune geometry is formed.}
\sym{m_{j,k}^{\mathrm{mov}},\,m_{j,k}^{\mathrm{fit}}}{The quantities $m_{j,k}^{\mathrm{mov}}$ and $m_{j,k}^{\mathrm{fit}}$ are the recent normalized motion and fit-activity indicators of block $j$, so they summarize whether that block has been moving or contributing enough to justify keeping it on the scaffold.}
\sym{\alpha_{\mathrm{pr}}}{The weight $\alpha_{\mathrm{pr}}\in[0,1]$ mixes motion-based and fit-based stagnation evidence, so it determines whether the prune gate leans more heavily on coordinate inactivity or on weak contribution to drift fitting.}
\sym{N_{\mathrm{prot}}}{The integer $N_{\mathrm{prot}}$ is the protection age for scaffold blocks, so it prevents freshly appended structure from being deleted before it has had enough checkpoint time to prove whether it is useful.}
\sym{\tau_{\mathrm{stale}},\,\tau_{\mathrm{loss}}^{\mathrm{pr}},\,\tau_{\theta}^{\mathrm{pr}},\,\tau_{\psi}^{\mathrm{pr}},\,\tau_{\rho}^{\mathrm{pr}}}{These thresholds are the prune gates for stagnation, cached projective loss, deleted-block norm, exact reduced-state jump, and reduced miss increase, so together they enforce that deletion is allowed only when a block is both old and truly inert in multiple geometric senses.}
\sym{\sigma_{j,k}^{\mathrm{stag}}}{The stagnation score $\sigma_{j,k}^{\mathrm{stag}}$ is the mixed inactivity measure of block $j$ at checkpoint $k$, so it is the first filter that turns recent coordinate behavior into a quantitative stale-enough criterion for prune eligibility.}
\sym{\operatorname{age}_{j,k},\,\operatorname{cooldown}_{j,k}}{The metadata $\operatorname{age}_{j,k}$ and $\operatorname{cooldown}_{j,k}$ are the age and refractory counters of block $j$, so they prevent premature or repeatedly failed deletions from being reconsidered as though nothing had happened.}
\sym{\mathcal D_k^{\mathrm{pr}}}{The set $\mathcal D_k^{\mathrm{pr}}$ is the set of prune-eligible blocks at checkpoint $k$, so it is the calm-regime deletion menu over which the exact local prune law searches.}
\sym{G_{k,-j},\,f_{k,-j}}{The reduced geometry $G_{k,-j}$ and reduced force $f_{k,-j}$ are the metric and force after deleting the coordinates belonging to block $j$, so they are the exact fixed-scaffold ingredients of the hypothetical reduced system obtained if that block were removed.}
\sym{\dot\theta_{k,-j}^{\mathrm{proj}}}{The reduced projected velocity $\dot\theta_{k,-j}^{\mathrm{proj}}:=G_{k,-j}^{+}f_{k,-j}$ is the unregularized McLachlan solution on the scaffold with block $j$ deleted, so it tells the prune law what the locally best reduced tangent motion would be if that deletion were accepted.}
\sym{L_{j,k}^{\mathrm{cache}}}{The cached prune loss $L_{j,k}^{\mathrm{cache}}$ is the normalized loss in projective drift-capture score caused by deleting block $j$ according to cached geometry only, so it is the cheap first-pass estimate of how much expressive power would be sacrificed by the deletion.}
\sym{|\psi_k^{(-j)}\rangle}{The reduced state $|\psi_k^{(-j)}\rangle$ is the actual variational state on the deleted scaffold obtained after removing block $j$, so it is the authoritative state used to verify that the cached prune loss did not hide a real physical jump.}
\sym{\theta_k^{(-j)},\,\mathcal O_k^{(-j)}}{The reduced coordinates $\theta_k^{(-j)}$ and reduced scaffold $\mathcal O_k^{(-j)}$ are the checkpoint state data after deleting block $j$, so they identify the precise variational configuration whose state and miss ratio are used in the final prune accept test.}
\sym{\Theta_{j,k}^{\mathrm{blk}}}{The block norm $\Theta_{j,k}^{\mathrm{blk}}$ is the Euclidean norm of the coordinates owned by block $j$, so it tests whether the block is already small enough in amplitude that deleting it is unlikely to cause a meaningful coordinate-space discontinuity.}
\sym{\Delta\psi_{j,k}^{\mathrm{pr}}}{The reduced-state jump $\Delta\psi_{j,k}^{\mathrm{pr}}$ is the Hilbert-space norm difference between the current state and the deleted-scaffold state, so it is the state-level safeguard that prevents deleting a numerically quiet block that nonetheless changes the physical state too much.}
\sym{E_{\psi_k^{(-j)},k}^{\sharp}}{The reduced-state sampled energy $E_{\psi_k^{(-j)},k}^{\sharp}$ is the sampled Hamiltonian expectation in the deleted-scaffold state, so it centers the reduced drift and makes the reduced miss ratio compare like with like against the current scaffold.}
\sym{\rho_{\mathrm{miss},k}^{(-j)}}{The reduced miss ratio $\rho_{\mathrm{miss},k}^{(-j)}$ is the normalized projective defect of the deleted scaffold, so it tells the prune law how much geometric inadequacy would remain if block $j$ were actually removed.}
\sym{\Delta\rho_{j,k}^{\mathrm{pr}}}{The miss increase $\Delta\rho_{j,k}^{\mathrm{pr}}$ is the change in normalized projective defect caused by deleting block $j$, so it is the final geometric safeguard against pruning a block whose removal makes the scaffold too underexpressive.}
\sym{\operatorname{accept\_prune}(j)}{The predicate $\operatorname{accept\_prune}(j)$ is the exact local prune accept condition for block $j$, so it says that deletion is allowed only when cached loss, coordinate amplitude, state jump, and miss-ratio increase are all simultaneously small.}

\subsection*{17A.10 Canonical piecewise-smooth law}

\sym{\mathcal O(\tau)}{The time-dependent scaffold $\mathcal O(\tau)$ is the piecewise-constant ordered block list active at physical time $\tau$, so it is the structural part of the law that changes only at discrete checkpoint events while remaining frozen during continuous projected evolution.}
\sym{\theta(\tau)}{The time-dependent coordinate vector $\theta(\tau)$ is the continuous parameter path on the currently active scaffold, so it carries the smooth between-checkpoint McLachlan motion once the discrete scaffold has been fixed for that interval.}
\sym{\tau\text{ in 17A.10}}{The symbol $\tau$ is the continuous physical time variable of the piecewise-smooth law, so it is the real-time parameter with respect to which the state evolves continuously between checkpoint events.}
\sym{\tau_0<\tau_1<\tau_2<\cdots}{The ordered sequence $\tau_0<\tau_1<\tau_2<\cdots$ is the checkpoint schedule, so it marks the discrete event times at which the law is allowed to switch scaffolds, recompute geometry, and choose stay, append, or prune.}
\sym{|\psi(\tau)\rangle}{The time-dependent state $|\psi(\tau)\rangle$ is the variational trajectory obtained by evaluating the active scaffold at the evolving coordinates, so it is the final piecewise-smooth quantum path produced by the whole continuous-plus-discrete control law.}
\sym{(\mathcal O_{k+1},\theta_{k+1})}{The pair $(\mathcal O_{k+1},\theta_{k+1})$ is the post-checkpoint scaffold and coordinate data handed to the next interval, so it is the discrete output of the checkpoint controller after the forecast or prune rule has been applied.}
\sym{\theta_{k+1}^{\mathrm{stay}}}{The coordinates $\theta_{k+1}^{\mathrm{stay}}$ are the next-interval coordinates when the controller keeps the scaffold unchanged, so they are the checkpoint output of a successful stay action.}
\sym{r_k^{\star}}{The record $r_k^{\star}$ is the winning append record when scaffold growth is selected at checkpoint $k$, so it identifies exactly which candidate block and insertion position are committed to the scaffold.}
\sym{a_k^{\star}}{The scale $a_k^{\star}$ is the winning discrete forecast scale attached to the selected append action, so it tells the next interval how strongly the committed append proposal was extrapolated when it won the forecast comparison.}
\sym{\vartheta_{k+1}^{\mathrm{app}}}{The coordinates $\vartheta_{k+1}^{\mathrm{app}}$ are the next-interval coordinates on the enlarged scaffold after a winning append action, so they are the checkpoint output that seeds continuous projected motion on the newly grown variational manifold.}
\sym{j_k^{\star}}{The index $j_k^{\star}$ is the winning block selected by the calm-regime prune law, so it identifies exactly which existing scaffold block is deleted when the controller chooses contraction instead of stay or growth.}
\sym{\theta_{k+1}^{(-j_k^{\star})}}{The coordinates $\theta_{k+1}^{(-j_k^{\star})}$ are the next-interval coordinates on the scaffold with block $j_k^{\star}$ deleted, so they are the post-prune state data that launches the next interval after a successful calm-regime contraction.}

\end{document}
